---
title: "US Census Data for Jackson County"
author: "Matthew Moore"
date: "4/22/2020"
output: html_document
---
## What does the U.S. Census data show us about the income statistics for Jackson County, where Newport School District resides?
Load libraries and data
```{r}
library(tidyverse)
library(rio)
library(janitor)
library(knitr)

#Import Data
ArkCensus <- rio::import("https://github.com/mrm081/busdrivers/blob/master/USArk_Counties_Poverty_ACS_16_5YR_DP03_Jan_24.xlsx?raw=true", which = "Poverty")
```
Filter data
```{r}
RichPoor <- ArkCensus %>% 
  select(Geography, Median_Househ_Income) %>% 
  group_by(Geography) %>% 
  arrange(desc(Median_Househ_Income))

PoorestAR <- ArkCensus %>% 
  top_n(-11, Median_Househ_Income) %>%
  select(Geography, Pct_Income_to_25k, Median_Househ_Income) %>% 
  arrange(Median_Househ_Income)
```
According to our data, we see that Jackson County is the 11th poorest county in Arkansas.

Let's see the graph.
```{r}
PoorestAR %>% 
  ggplot(aes(x = reorder(Geography, -Median_Househ_Income), y = Median_Househ_Income, fill = Median_Househ_Income)) +
  geom_col(position = "dodge", show.legend = FALSE) +
  theme(axis.text.x = element_text(angle = 90)) +
  geom_text(aes(label = Median_Househ_Income), hjust = -.25, vjust = 0, size = 3.75) +
  expand_limits(x = 12, y = 35000) +
  coord_flip() +
  labs(title = "Poorest Counties in Arkansas",
       subtitle = "2016 U.S. Census Data",
       caption = "Graphic by Matthew Moore - 4/22/20",
       y="Median Household Income",
       x="")
```
Let's look at a few select counties in Arkansas: Washington and Benton in Northwest Arkansas, Pulaski and Faulkner in Central Arkansas, and let's also include the state overall and graph it.

```{r}
Select_Counties <- ArkCensus %>% 
  filter(Geography=="Arkansas"|Geography=="Benton County, Arkansas"|Geography=="Washington County, Arkansas"|
           Geography=="Jackson County, Arkansas" | Geography=="Pulaski County, Arkansas" |
           Geography=="Faulkner County, Arkansas") %>% 
  select(Geography, Pct_Income_to_25k, Median_Househ_Income)

Select_Counties %>% 
  ggplot(aes(x = reorder(Geography, -Median_Househ_Income), y = Median_Househ_Income, fill = Median_Househ_Income)) +
  geom_col(position = "dodge", show.legend = FALSE) +
  theme(axis.text.x = element_text(angle = 90)) +
  geom_text(aes(label = Median_Househ_Income), hjust = -.25, vjust = 0, size = 3.75) +
  expand_limits(x = 6, y = 65000) +
  coord_flip() +
  labs(title = "Select Counties in Arkansas",
       subtitle = "2016 U.S. Census Data",
       caption = "Graphic by Matthew Moore - 4/22/20",
       y="Median Household Income",
       x="")
```
