---
title: "Newport vs. State Homeless Data"
author: "Matthew Moore"
date: "4/21/2020"
output: html_document
---

## How does Newport School District compare to the State of Arkansas overall when it comes to homelessness? Let's find out.

Install libraries
```{r include=FALSE}
library(tidyverse)
library(rio)
library(janitor)
library(formattable)
```
Import and filter 2018/2019 Data
```{r}
zData1819 <- rio::import("https://github.com/mrm081/busdrivers/blob/master/AR_School_Homeless_Data1819.xlsx?raw=true", which = "Data")
zData1819 <- janitor::clean_names(zData1819)
zData1819[3:16] <- lapply(zData1819[3:16], as.numeric)
Schools1819 <- zData1819 %>% 
  select(district_name, district_percent_homeless) %>% 
  filter(district_name == "Arkansas Overall" | district_name == "NEWPORT SCHOOL DISTRICT")
colnames(Schools1819)[2] <- c("homeless_pct_1819")
head(Schools1819)
```
Import and Filter 2014/2015 Data
```{r}
zData1415 <- rio::import("https://github.com/mrm081/busdrivers/blob/master/AR_School_Homeless_data1415.xlsx?raw=true", which = "Data")
zData1415 <- janitor::clean_names(zData1415)
zData1415[3:16] <- lapply(zData1415[3:16], as.numeric)
Schools1415 <- zData1415 %>% 
  select(district_name, district_percent_homeless) %>% 
  filter(district_name == "Arkansas Overall" | district_name == "NEWPORT SCHOOL DISTRICT")
colnames(Schools1415)[2] <- c("homeless_pct_1415")
head(Schools1415)
```

Join Data together
```{r}
CompareData<- Schools1415 %>% 
  inner_join(Schools1819, by=("district_name"))
head(CompareData)
```
Find percentage change between 2014/2015 school year to 2018/2019 school year.
```{r}
CompareData <- CompareData %>% 
mutate(PctChg = (homeless_pct_1819 - homeless_pct_1415)/homeless_pct_1819) 
CompareData$PctChg <- percent(CompareData$PctChg)
CompareData$homeless_pct_1415 <- percent(CompareData$homeless_pct_1415)
CompareData$homeless_pct_1819 <- percent(CompareData$homeless_pct_1819)
head(CompareData)
```
Graphing the Percentage change

```{r}
CompareData %>% 
  ggplot(aes(x = district_name, y = PctChg, fill = PctChg)) +
geom_col(position = "dodge", show.legend = FALSE) +
  theme(axis.text.x = element_text(angle = 90)) +
geom_text(aes(label = PctChg), hjust = -.5, vjust = -.5, size = 3.5) +
  expand_limits(x = 2, y = .35) +
  coord_flip() +
  labs(title = "Percent Change in Homeless Students",
       subtitle = "14/15 School Year - 18/19 School Year",
       caption = "Graphic by Matthew Moore - 4/21/20",
       y="Percentage Change",
       x="School")
```

This chart is not very representative of the data, so let's try another couple of graphs.
```{r}
CompareData %>% 
  ggplot(aes(x = district_name, y = homeless_pct_1415, fill = homeless_pct_1415)) +
geom_col(position = "dodge", show.legend = FALSE) +
  theme(axis.text.x = element_text(angle = 90)) +
geom_text(aes(label = homeless_pct_1415), hjust = -.5, vjust = -.5, size = 3.5) +
  expand_limits(x = 2, y = .11) +
    coord_flip() +
  labs(title = "Homeless Students Percentage",
       subtitle = "2014/2015 School Year",
       caption = "Graphic by Matthew Moore - 4/21/20",
       y="Percentage Change",
       x="")
```

```{r}
CompareData %>% 
  ggplot(aes(x = district_name, y = homeless_pct_1819, fill = homeless_pct_1819)) +
geom_col(position = "dodge", show.legend = FALSE) +
  theme(axis.text.x = element_text(angle = 90)) +
geom_text(aes(label = homeless_pct_1819), hjust = -.5, vjust = -.5, size = 3.5) +
  expand_limits(x = 2, y = .13) +
    coord_flip() +
  labs(title = "Homeless Students Percentage",
       subtitle = "2018/2019 School Year",
       caption = "Graphic by Matthew Moore - 4/21/20",
       y="Percentage Change",
       x="")
```